Thorough example

### 2021-01-11 inst/gitbook/05_thorough_example.Rmd
Update 05_thorough_example.Rmd

 1. A lineage evolves in a normal birth-death fashion (with extinction) from a single origin and with two traits called:
    - "Distance": the physical distance between species in an arbitrary 1D space (evolving in a Brownian motion).
    - "Variable": a 2D uncorrelated Ornstein-Uhlenbeck trait that corresponds to some traits the lineages have.
 During the whole process, species that are more distant from their ancestor have a greater change of speciating.

@@

 2. "Correlation event": after reaching a certain number of taxa an event happens making the "Variable" trait dimensions become correlated.
 3. "Selective extinction event": after reaching a certain point in time, lineages with their "Distance" values in some interval goes extinct.
 4. "Founding event": after reaching a value for the "Distance" trait, a founding event happens with a new birth-death process with no extinction and where the "Variable" is now a 2D correlated Brownian Motion.
 5. "Mass extinction event": after reaching a certain number of taxa in the "Founding event", 80% of taxa go extinct randomly.

@@

```{r}
## The OU 2D trait correlation
correlation <- matrix(0, 2, 2)

## The trait object
two_traits <- make.traits(process = c(BM.process, OU.process), n = c(1, 2), 
                          process.args = list(list(NULL), list(Sigma = correlation)),
                          trait.names = c("Distance", "Variables"))
#TODO: fix OU with correlation
```

@@

## And displaying the results
par(mfrow = c(1,2))
plot(results, trait = 1, main = "Trait 1: Distance")
plot(results, trait = c(2,3), main = "Variables correlation")

## TODO: debug correlation!
```

@@

```{r}
## Make a distance modifier
distance_speciation <- make.modifiers(speciation = NULL,
                                      condition = NULL,
                                      modify = NULL) 
```

```{r}
## Running a birth death tree 
results <- dads(bd.params  = bd_params,
                stop.rule  = stop_rule,
                traits     = two_traits,
                modifiers  = distance_speciation,
                null.error = 100)

@@

```{r}
## The correlation change event
events_list <- make.events(event.name   = "Correlation",
                           target       = "traits",
                           condition    = taxa.condition(30),
                           modification = update.traits(
                    process.args = list(Sigma = matrix(c(0.7, 0, 0, 0.7), 2, 2)),
                    trait.names  = "Variables"))
#TODO: fix bug here
```

```{r}
## Running a birth death tree 
results <- dads(bd.params  = bd_params,
                stop.rule  = stop_rule,
                traits     = two_traits,
                modifiers  = distance_speciation,
                events     = events_list,
                null.error = 100)

## And displaying the results
par(mfrow = c(1,2))
plot(results, trait = 1, main = "Trait 1: Distance")
plot(results, trait = c(2,3), main = "Variables correlation")
## TODO: debug correlation!
```

## Step 3:
 3. "Selective extinction event": after reaching a certain point in time, lineages with their "Distance" values in some interval goes extinct.

```{r}
## Adding the selective event
events_list <- make.events(event.name   = "Selection",
                           add          = events_list,
                           target       = "taxa",
                           condition    = time.condition(4),
                           modification = trait.extinction(1, condition = `>`),
                           event.name   = "Selection")
```

```{r}
## Running a birth death tree 
results <- dads(bd.params  = bd_params,
                stop.rule  = stop_rule,
                traits     = two_traits,
                modifiers  = distance_speciation,
                events     = events_list,
                null.error = 100)

## And displaying the results
par(mfrow = c(1,2))
plot(results, trait = 1, main = "Trait 1: Distance")
plot(results, trait = c(2,3), main = "Variables correlation")
## TODO: debug correlation!
```

@@

```{r}
## The new trait for the founding event
new_traits <- make.traits(process = c(BM.process, BM.process), n = c(1, 2), 
                          process.args = list(
                              list(NULL),
                              list(Sigma = matrix(c(0.7, 0, 0, 0.7), 2, 2))),
                          trait.names = c("Distance", "Variables"))
#TODO: set starting value to be ancestors' value

@@

```{r}
## The selective event
events_list <- make.events(event.name   = "Founding", 
                           add          = events_list,
                           target       = "founding",
                           condition    = trait.condition(3),
                           modification = founding.event(traits    = new_traits,
                                                         events    = mass_extinction,
                                                         bd.params = list(
                                                            speciation = 2,
                                                            extinction = 0)))
#TODO: fix bug
```

```{r}
## Running a birth death tree 
results <- dads(bd.params  = bd_params,
                stop.rule  = stop_rule,
                traits     = two_traits,
                modifiers  = distance_speciation,
                events     = events_list,
                null.error = 100)

## And displaying the results
par(mfrow = c(1,2))
plot(results, trait = 1, main = "Trait 1: Distance")
plot(results, trait = c(2,3), main = "Variables correlation")
## TODO: debug correlation!
```

>>>>

 1. A lineage evolves in a normal birth-death fashion (with extinction) from a single origin and with two traits called:
    - "Distance": the physical distance between species in an arbitrary 1D space (evolving in a Brownian motion).
    - "Variable": a 2D uncorrelated Ornstein-Uhlenbeck trait that corresponds to some traits the lineages have.
 During the whole process, lineages that are more further away from the other species in terms of distance have a higher change of speciating (proportional to that distance).

@@

 2. "Correlation event": after reaching a certain number of taxa an event happens making the "Variables" trait dimensions become correlated.
 3. "Selective extinction event": after reaching a certain point in time, lineages with their "Variables" values that are negative go extinct
 4. "Founding event": after reaching a value for the "Distance" trait, a founding event happens with a new birth-death process with no extinction and where the "Variable" is now a 2D correlated Brownian Motion.
 5. "Mass extinction event": after reaching a certain number of taxa in the "Founding event", 80% of taxa go extinct randomly.

@@

```{r}
## The trait object
two_traits <- make.traits(process = c(BM.process, OU.process), n = c(1, 2),
                          trait.names = c("Distance", "Variables"))
```

```{r}
## Visualise the 1D trait (BM.process)
plot(two_traits, trait = 1)
```

```{r, eval = FALSE}
## Visualise the 2D trait (uncorrelated OU.process)
plot(two_traits, trait = 2, use.3D = TRUE)
```

@@

@@

```{r}
## Speciation event is more likely if lineage's ancestor is further away from the mean trait value
distance.modify <- function(x, trait.values, lineage) return(x + x * abs(parent.traits(trait.values, lineage)[1] - mean(trait.values[,1])))

## Make a distance modifier
distance.speciation <- make.modifiers(speciation = speciation,
                                      modify = distance.modify)
```

```{r}
## Running a birth death tree 
results <- dads(bd.params  = bd_params,
                stop.rule  = stop_rule,
                traits     = two_traits,
                modifiers  = distance.speciation,
                null.error = 100)

@@

```{r}
## The correlation change event
events_list <- make.events(event.name   = "Correlation",
                           target       = "traits",
                           condition    = taxa.condition(30),
                           modification = update.traits(
                    process.args = list(Sigma = matrix(c(0.7, 0, 0, 0.7), 2, 2)),
                    trait.names  = "Variables"))
```

```{r}
## Running a birth death tree 
results <- dads(bd.params  = bd_params,
                stop.rule  = stop_rule,
                traits     = two_traits,
                modifiers  = distance.speciation,
                events     = events_list,
                null.error = 100)

## Step 3:
 3. "Selective extinction event": after reaching a certain point in time, lineages with their "Variables" values that are negative go extinct

```{r}
## Adding the selective event
events_list <- make.events(event.name   = "Selection",
                           add          = events_list,
                           target       = "taxa",
                           condition    = time.condition(4),
                           modification = trait.extinction(0, condition = `<`, trait = 2))
```

```{r}
## Running a birth death tree 
results <- dads(bd.params  = bd_params,
                stop.rule  = stop_rule,
                traits     = two_traits,
                modifiers  = distance.speciation,
                events     = events_list,
                null.error = 100)

@@

```{r}
## The new trait for the founding event
new_traits <- make.traits(process = c(BM.process, BM.process), n = c(1, 2),
                          trait.names = c("Distance", "Variables"))

@@

```{r}
## The selective event
events_list <- make.events(event.name   = "Founding", 
                           add          = events_list,
                           target       = "founding",
                           condition    = trait.condition(3),
                           modification = founding.event(traits    = new_traits,
                                                         events    = mass_extinction,
                                                         bd.params = list(
                                                            speciation = 2,
                                                            extinction = 0)),
                           additional.args = list(prefix = "found_"))
```

```{r}
## Running a birth death tree 
results <- dads(bd.params  = bd_params,
                stop.rule  = stop_rule,
                traits     = two_traits,
                modifiers  = distance.speciation,
                events     = events_list,
                null.error = 100)
